Finch: Sparse and Structured Array Programming with Control Flow

### 2024-04-06 sections/eval/graphs.tex
fixes
separate

Figure~\ref{fig:graph_result} compares performance to Graphs.jl, a Julia library, and the LAGraph Library, which implements graph algorithms with sparse linear algebra using GraphBLAS~\cite{mattson2019lagraph}.
%
For the BFS algorithm, direction-optimization notably enhances performance for scale-free graphs. 
%
Although GraphBLAS uses hardwired optimizations, Finch is competitive. 
%
On Bellman-Ford (with path lengths and shortest-path tree), Finch's support for multiple outputs, sparse inputs, and masks leads to superior performance over GraphBLAS (average speedup of 3.53). 
%
We did not include GAP-road as it timed out. %speedup is arith-mean
%
In Appendix~\ref{sec:graph_listings}, we display the code for BFS and Bellman-Ford in Finch (57 and 50 LOC) and LAGraph (215 and 227 LOC), and invite readers to compare the clarity of the algorithms.

\begin{figure}
	\includegraphics[width=0.5\linewidth]{bfs_speedup_over_graphs.jl.png}%
	\includegraphics[width=0.5\linewidth]{bellmanford_speedup_over_graphs.jl.png}
    \vspace{-12pt}
    \caption{Performance of graph apps across various tools. finch\_push\_only exclusively utilizes push traversal, while finch\_push\_pull applies direction-optimization akin to GraphBLAS. Finch's support for push/pull traversal and early break facilitates direction-optimization. Among GraphBLAS's five variants for Bellman-Ford, we selected LAGraph\_BF\_full1a, consistently the fastest with our graphs.}
     \label{fig:graph_result}
\end{figure}

\begin{figure}
    \begin{minipage}{0.33\linewidth}
    \begin{minted}{julia}
    V = Tensor(Dense(Element(false)))
    P = Tensor(Dense(Element(0)))
    F = Tensor(SparseByteMap(Pattern()))
    _F = Tensor(SparseByteMap(Pattern()))
    A = Tensor(Dense(SparseList(Pattern())))
    AT = Tensor(Dense(SparseList(Pattern())))

@@

>>>>

Figure~\ref{fig:graph_result} compares performance to Graphs.jl, a Julia library, and the LAGraph Library, which implements graph algorithms with sparse linear algebra using GraphBLAS~\cite{mattson2019lagraph}.
%
For the BFS algorithm, direction-optimization notably enhances performance for scale-free graphs. 
%
Although GraphBLAS uses hardwired optimizations, Finch is competitive. 
%
On Bellman-Ford (with path lengths and shortest-path tree), Finch's support for multiple outputs, sparse inputs, and masks leads to superior performance over GraphBLAS (average speedup of 3.53). 
%
We did not include GAP-road as it timed out. %speedup is arith-mean
%
In Appendix B, we display the code for BFS and Bellman-Ford in Finch (57 and 50 LOC) and LAGraph (215 and 227 LOC), and invite readers to compare the clarity of the algorithms.

\begin{figure}[b]
    \begin{minipage}{0.33\linewidth}
    \begin{minted}{julia}
    V = Tensor(Dense(Element(false)))
    P = Tensor(Dense(Element(0)))
    F = Tensor(SparseByteMap(Pattern()))
    _F = Tensor(SparseByteMap(Pattern()))
    A = Tensor(Dense(SparseList(Pattern())))
    AT = Tensor(Dense(SparseList(Pattern())))

@@

\begin{figure}[t]
	\includegraphics[width=0.5\linewidth]{bfs_speedup_over_graphs.jl.png}%
	\includegraphics[width=0.5\linewidth]{bellmanford_speedup_over_graphs.jl.png}
    \vspace{-12pt}
    \caption{Performance of graph apps across various tools. finch\_push\_only exclusively utilizes push traversal, while finch\_push\_pull applies direction-optimization akin to GraphBLAS. Finch's support for push/pull traversal and early break facilitates direction-optimization. Among GraphBLAS's five variants for Bellman-Ford, we selected LAGraph\_BF\_full1a, consistently the fastest with our graphs.}
     \label{fig:graph_result}
\end{figure}

### 2024-04-06 sections/eval/morphology.tex
separate

We compared against OpenCV.
%
We used four datasets. We randomly selected 100 images from the MNIST \cite{lecun_gradient-based_1998} and Omniglot \cite{lake_human-level_2015} character recognition datasets, as well as a dataset of human line drawings \cite{eitz_how_2012}. 
%
We also hand-selected a subset of mask images (these images were less homogenous, so we listed them in the appendix) from a digital image processing textbook \cite{gonzalez_digital_2006}. 
%
All images were thresholded, and we also include versions of the images that have been magnified before thresholding, to induce larger constant regions. 
%
IN our erosion task, the sparseRLE format performs the best as it is asymptotically faster and uses less memory, leading to a 19.5X speedup over OpenCV on the sketches dataset, which becomes arbitrarily large as we magnify the images (here shown as 266X). 
%
We believe the 51.6x on MNIST is due to calling overhead in OpenCV). 
%
The bitwise kernels were effective as well, and would be more effective on datasets with less structure. 
%
A strength of Finch is that it supports structured datasets, even over bitwise operations, allowing us to implement the bitwise kernel and then mask it.

>>>>

We compared against OpenCV.
%
We used four datasets. We randomly selected 100 images from the MNIST \cite{lecun_gradient-based_1998} and Omniglot \cite{lake_human-level_2015} character recognition datasets, as well as a dataset of human line drawings \cite{eitz_how_2012}. 
%
We also hand-selected a subset of mask images (these images were less homogenous, so we listed them in Appendix C) from a digital image processing textbook \cite{gonzalez_digital_2006}. 
%
All images were thresholded, and we also include versions of the images that have been magnified before thresholding, to induce larger constant regions. 
%
IN our erosion task, the sparseRLE format performs the best as it is asymptotically faster and uses less memory, leading to a 19.5X speedup over OpenCV on the sketches dataset, which becomes arbitrarily large as we magnify the images (here shown as 266X). 
%
We believe the 51.6x on MNIST is due to calling overhead in OpenCV). 
%
The bitwise kernels were effective as well, and would be more effective on datasets with less structure. 
%
A strength of Finch is that it supports structured datasets, even over bitwise operations, allowing us to implement the bitwise kernel and then mask it.
